Supporting Information : Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes

### 2018-10-08 Ben_Manuscripts/structure_paper/supporting_information.tex
Edits

  We chose to build all systems using 20 monomers per column, i.e. for a simulation with 5 columns around each pore, there are 100 monomers per pore, and 400 monomers per simulation.  We made this choice
  by considering a number of factors that directly affect the quality of our results. 
  Clearly, choosing a smaller number of layers will cause the system to suffer from 
  larger finite size effects. Even in our 20 monomer-per-column system, axial correlations
  between stacked monomers persist throughout the entire correlation function (see
  Figure~\ref{M-fig:correlation} of the main text). The correlation function of the system which we
  built with 40 monomers per column is very similar to that created with 20 monomers
  per columns (continues to oscillate for its full length
  (Figure~\ref{fig:tall_correlation_full})). If we continue adding monomers, the 
  correlation function will eventually fully decay but it is not worth the 
  computational expense. We do find that the correlation lengths calculated from 20
  monomer per column simulations are in sufficient agreement with experiment. 
  %MRSF: if 10 monomers per column gave inconsistent pores when tried, might as well mention that here. 
  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{z_correlation_overlay.png}
  \caption{}\label{fig:z_correlation_overlay}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{tall_correlation_full.png}
  \caption{}\label{fig:tall_correlation_full}
  \end{subfigure}
  \caption{(a) The correlation functions generated from ordered basin sandwiched 
  configurations are nearly the same when the $z$-dimension of the system is doubled.
  (b) Oscillations persist throughout the length of the correlation function for the
  sandwiched system in the ordered basin built with 40 monomers per column. The full
  system is ca. 17 nm tall so the full correlation function covers half this length
  since periodicity forces the remaining length to be its mirror image.}\label{fig:correlation_functions}
  \end{figure}

@@

 %BJC: is it really necessary to explain this if I release the code? I don't explain
 %any of the other stats I calculate. 
  We are interested in 5 pore-to-pore distances which should all be
  equal in a perfect hexagonal array (Figure~\ref{fig:p2p_diagram}). Only 4 of
  the 5 distances are independent. We can calculate a trajectory of spacing
  versus time for each of the 5 distances. We calculated the average pore-to-pore
  spacing and its uncertainty according to the following procedure:
   \begin{enumerate}
	\item We calculated the time when each of the pore-to-pore distances
	were equilibrated using \\ \texttt{pymbar.timeseries.detectEquilibration()}
	\cite{chodera_simple_2016,shirts_statistically_2008}. We began calculations
	after the largest of the five values.
	\item We calculated how long it takes for the data in each of the 5
	trajectories to become uncorrelated using \texttt{pymbar.timeseries.integratedAutocorrelationTime()}
	\cite{chodera_simple_2016,shirts_statistically_2008}. 
	\item We broke the full equilibrated trajectory into blocks of length $\tau$, where
	$\tau$ is the maximum of the five autocorrelation times calculated. Each block contains
	five sub-trajectories of pore-to-pore spacings.   
	\item We generate statistics using the bootstrapping technique. For
	each bootstrap trial, we reconstruct an equilibrium trajectory by randomly
	sampling from the trajectory blocks. 
	\item The average pore-to-pore distance is the mean of all pore
	spacings among all bootstrap trials.
	\item To calculate the uncertainty in pore-to-pore distance, we calculate the average 
	pore spacing for each pore over all bootstrap trials. Using the 5 average 
	pore-to-pore distances, we calculate the spread with:
	\begin{equation}
	s = \sqrt{\frac{1}{4} \sum_{i=1}^5 (x_i - \overline{x})^2} 
	\label{equation:p2p_stats}
	\end{equation}
	where $\overline{x}$ is the average pore-to-pore distance. We divided the sum in 
	equation~\ref{equation:p2p_stats} by 4 since there are only 4 independent pore
	spacings.
        %MRSF: I think this is the spread in a single pore, not the average pore.  Should probably make it clearer to readers which it is (i.e. uncertainty in one, not uncertainty in mean of the four).
  \end{enumerate}
  \begin{figure}[!htb]
	\centering
	\includegraphics[width=0.25\textwidth]{p2p_diagram.png}
	\caption{There are five pore-to-pore distances that are equal in a perfect
		hexagon. If each number in the diagram represents the center of a
		pore in a hexagonal unit cell, then the distance from 1 to 2, 2 to 4, 
		4 to 3, 3 to 1 and 1 to 4 should be equal. Only 4 of the distances
		are independent. For example, the distance from 1 to 4 is defined by
		the location of all other pore centers.}
	\label{fig:p2p_diagram}  %MRSF: this label is multiply defined. 
  \end{figure}

@@

  In the main text, we calculated ionic conductivity using the Nernst-Einstein 
  relationship. We also measured ionic conductivity using a second method, 
  %MRSF: Don't capitalize Collective Diffusion. Also, add citation on first mention.
  termed the Collective Diffusion model, for robustness. The collective 
  diffusion model measures the movement of the collective variable, Q, which
  is defined as the amount of charge transfer through the system and can be 
  thought to represent the center of charge of the system. The conductance, 
  $\gamma$, of the system can be calculated as:
  \begin{equation}
	 \gamma = \dfrac{D_Q}{k_b T} 
	\label{eqn:collective_diffusion}
  \end{equation}
  $D_Q$ is the diffusion coefficient of the collective variable Q and is 
  calculated using the Einstein relation. We convert the resulting value
  to ionic conductivity by multiplying by channel length and dividing by
  the membrane cross sectional area. One can access a detailed derivation of 
  the model elsewhere\cite{liu_collective_2013}.

@@

  \section{Ensemble of ordered sandwiched configurations}

@@

  %MRSF: isn't this picture the same in the main draft of the document. Probably wouldn't have occured.
  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{sandwiched_ensemble_stds.pdf}
  \caption{}\label{fig:offset_ensemble_stds}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{sandwiched_ensemble_means.pdf}
  \caption{}\label{fig:offset_ensemble_means}
  \end{subfigure}
  \begin{subfigure}{\textwidth}
  \includegraphics[width=\textwidth]{sandwiched_ensemble_regional_density.pdf}
  \caption{}\label{fig:offset_ensemble_regional_density}
  \end{subfigure}
  \caption{(a) The standard deviation of the distribution of quenched disorder
  from the first 5 ns of the main ordered sandwiched system studied in
  this paper (black dashed line) is in agreement with the distribution of quenched disorder 
  standard deviations calculated from the ensemble of simulations (histogram). 
  (b) The mean values of $r$ and $\theta$ from the first 5 ns of the main
  system trajectory (black dashed line) is in agreement with the distribution of mean values
  calculated from each simulation in the ensemble (histogram). The mean values of $z$
  are necessarily 0 so they are not plotted. (c) The radial densities of tail
  atoms, head group atoms and sodium atoms calculated from the first 5 ns of the
  main system simulation trajectory (black lines) and from the ensemble of 
  trajectories (all other lines) look qualitatively similar.}\label{fig:ensemble_stds}
  \end{figure}

@@

  %MRSF: probably needs some context, since this is for a force field, not for QM. Did you have plots for several different force fields, which all are presumably trying to capture the QM that all basically gave the same results? That would be better evidence.
  \begin{figure}[!htb]
  \centering
  \includegraphics[width=0.5\textwidth]{carboxylate_dihedral_rb.png}
  \caption{A potential energy diagram for the dihedral angle about the carbon-carbon bond of
  a carboxylate group attached to an aromatic ring, using parameters from the generalized amber 
  forcefield. It is energetically favorable for the carboxylate group to stay in plane with 
  the aromatic ring. There is a large energy penalty as the dihedral becomes antiparallel to 
  the plane of the aromatic ring.}\label{fig:carboxylate_dihedral_rb}
  \end{figure}

% BJC: not necessary
%  \begin{figure}[!htb]
%  \centering
%        \begin{subfigure}{0.45\textwidth}
%                \includegraphics[width=\textwidth]{water_filled_pore.png}
%                \caption{}\label{fig:water_filled_pores}
%        \end{subfigure}
%        \begin{subfigure}{0.45\textwidth}
%                \includegraphics[width=\textwidth]{water_removed.png}
%                \caption{}\label{fig:water_removed}
%        \end{subfigure}
%  \caption{(a) Pores built in the parallel displaced configuration with 5
%	  monomers per layer are filled with 5 wt\% water. (b) The same system is
%	  visualized with water molecules and sodium ions removed. Head groups vacate the
%	  pore region leaving an aqueous solution of water and sodium ions.}\label{fig:water_pores}
%  \end{figure}

@@

  \section{The role of water in the appearance of R-double}\label{section:rdouble}

>>>>

  We chose to build all systems using 20 monomers per column, i.e. for a simulation with 5 columns around each pore, there are 100 monomers per pore, and 400 monomers per simulation.  We made this choice
  by considering a number of factors that directly affect the quality of our results. 
  Clearly, choosing a smaller number of layers will cause the system to suffer from 
  larger finite size effects. Even in our 20 monomer-per-column system, axial correlations
  between stacked monomers persist throughout the entire correlation function (see
  Figure~\ref{M-fig:correlation} of the main text). The correlation function of the system which we
  built with 40 monomers per column is very similar to that created with 20 monomers
  per columns (continues to oscillate for its full length
  (Figure~\ref{fig:tall_correlation_full})). If we continue adding monomers, the 
  correlation function will eventually fully decay but it is not worth the 
  computational expense. We do find that the correlation lengths calculated from 20
  monomer per column simulations are in sufficient agreement with experiment.

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{z_correlation_overlay.png}
  \caption{}\label{fig:z_correlation_overlay}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{tall_correlation_full.png}
  \caption{}\label{fig:tall_correlation_full}
  \end{subfigure}
  \caption{(a) The correlation functions generated from ordered basin sandwiched 
  configurations are nearly the same when the $z$-dimension of the system is doubled.
  (b) Oscillations persist throughout the length of the correlation function for the
  sandwiched system in the ordered basin built with 40 monomers per column. The full
  system is ca. 17 nm tall so the full correlation function covers half this length
  since periodicity forces the remaining length to be its mirror image.}\label{fig:correlation_functions}
  \end{figure}

@@

 %BJC: is it really necessary to explain this if I release the code? I don't explain
 %any of the other stats I calculate. 
  We are interested in 5 pore-to-pore distances which should all be
  equal in a perfect hexagonal array (Figure~\ref{fig:p2p_diagram}). Only 4 of
  the 5 distances are independent. We can calculate a trajectory of spacing
  versus time for each of the 5 distances. We calculated the average pore-to-pore
  spacing and its uncertainty according to the following procedure:
   \begin{enumerate}
	\item We calculated the time when each of the pore-to-pore distances
	were equilibrated using \\ \texttt{pymbar.timeseries.detectEquilibration()}
	\cite{chodera_simple_2016,shirts_statistically_2008}. We began calculations
	after the largest of the five values.
	\item We calculated how long it takes for the data in each of the 5
	trajectories to become uncorrelated using \texttt{pymbar.timeseries.integratedAutocorrelationTime()}
	\cite{chodera_simple_2016,shirts_statistically_2008}. 
	\item We broke the full equilibrated trajectory into blocks of length $\tau$, where
	$\tau$ is the maximum of the five autocorrelation times calculated. Each block contains
	five sub-trajectories of pore-to-pore spacings.   
	\item We generate statistics using the bootstrapping technique. For
	each bootstrap trial, we reconstruct an equilibrium trajectory by randomly
	sampling from the trajectory blocks. 
	\item The average pore-to-pore distance is the mean of all pore
	spacings among all bootstrap trials.
	\item To calculate the uncertainty in pore-to-pore distance, we calculate the average 
	pore spacing for each pore over all bootstrap trials. Using the 5 average 
	pore-to-pore distances, we calculate the spread in a single pore-to-pore distance 
	with:
	\begin{equation}
	s = \sqrt{\frac{1}{4} \sum_{i=1}^5 (x_i - \overline{x})^2} 
	\label{equation:p2p_stats}
	\end{equation}
	where $\overline{x}$ is the average pore-to-pore distance. We divided the sum in 
	equation~\ref{equation:p2p_stats} by 4 since there are only 4 independent pore
	spacings.
  \end{enumerate}

%BJC: included in a figure above
%  \begin{figure}[!htb]
%	\centering
%	\includegraphics[width=0.25\textwidth]{p2p_diagram.png}
%	\caption{There are five pore-to-pore distances that are equal in a perfect
%		hexagon. If each number in the diagram represents the center of a
%		pore in a hexagonal unit cell, then the distance from 1 to 2, 2 to 4, 
%		4 to 3, 3 to 1 and 1 to 4 should be equal. Only 4 of the distances
%		are independent. For example, the distance from 1 to 4 is defined by
%		the location of all other pore centers.}
%	\label{fig:p2p_diagram}  
%  \end{figure}

@@

  In the main text, we calculated ionic conductivity using the Nernst-Einstein 
  relationship. We also measured ionic conductivity using a second method, 
  termed the collective diffusion model\cite{liu_collective_2013}, for robustness. The collective 
  diffusion model measures the movement of the collective variable, Q, which
  is defined as the amount of charge transfer through the system and can be 
  thought to represent the center of charge of the system. The conductance, 
  $\gamma$, of the system can be calculated as:
  \begin{equation}
	 \gamma = \dfrac{D_Q}{k_b T} 
	\label{eqn:collective_diffusion}
  \end{equation}
  $D_Q$ is the diffusion coefficient of the collective variable Q and is 
  calculated using the Einstein relation. We convert the resulting value
  to ionic conductivity by multiplying by channel length and dividing by
  the membrane cross sectional area. One can access a detailed derivation of 
  the model elsewhere\cite{liu_collective_2013}.

@@

  \section{Ensemble of ordered sandwiched configurations}\label{section:sandwiched_ensemble}

@@

  %MRSF: isn't this picture the same in the main draft of the document. Probably wouldn't have occured.
  %BJC: it's not identical. The main text uses the parallel displaced data.
  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{sandwiched_ensemble_stds.pdf}
  \caption{}\label{fig:offset_ensemble_stds}
  \end{subfigure}
  \begin{subfigure}{0.49\textwidth}
  \includegraphics[width=\textwidth]{sandwiched_ensemble_means.pdf}
  \caption{}\label{fig:offset_ensemble_means}
  \end{subfigure}
  \begin{subfigure}{\textwidth}
  \includegraphics[width=\textwidth]{sandwiched_ensemble_regional_density.pdf}
  \caption{}\label{fig:offset_ensemble_regional_density}
  \end{subfigure}
  \caption{(a) The standard deviation of the distribution of quenched disorder
  from the first 5 ns of the main ordered sandwiched system studied in
  this paper (black dashed line) is in agreement with the distribution of quenched disorder 
  standard deviations calculated from the ensemble of simulations (histogram). 
  (b) The mean values of $r$ and $\theta$ from the first 5 ns of the main
  system trajectory (black dashed line) is in agreement with the distribution of mean values
  calculated from each simulation in the ensemble (histogram). The mean values of $z$
  are necessarily 0 so they are not plotted. (c) The radial densities of tail
  atoms, head group atoms and sodium atoms calculated from the first 5 ns of the
  main system simulation trajectory (black lines) and from the ensemble of 
  trajectories (all other lines) look qualitatively similar.}\label{fig:ensemble_stds}
  \end{figure}

@@

  It is most energetically favorable for the carboxylate group attached to the monomer
  head groups to stay coplanar with the phenyl group to which it is attached
  (Figure~\ref{fig:carboxylate_dihedral_rb}). Although GAFF may over-predict the penalty
  for deviating from its minimum energy configuration relative to those calculated 
  by Rakitin and Pack~\cite{rakitin_necessity_2005} and Nelson and
  Borkman~\cite{nelson_internal_1998}, they all show that there is an appreciable barrier
  to rotation about the carboxylate-phenyl bond. The potential energy reaches its maximum
  at rotation angles corresponding to a carboxylate group that is antiparallel to the 
  plane of the phenyl ring to which it is attached (90\degree and 270\degree).

  \begin{figure}[!htb]
  \centering
  \includegraphics[width=0.5\textwidth]{carboxylate_dihedrals.pdf}
  \caption{A potential energy diagram for the dihedral angle about the carbon-carbon bond of
  a carboxylate group attached to an aromatic ring, using parameters from GAFF (blue), 
  Rakitin (orange) and Borkman (green). It is energetically favorable for the 
  carboxylate group to stay in plane with the aromatic ring. There is a large energy 
  penalty as the dihedral becomes antiparallel to the plane of the aromatic 
  ring.}\label{fig:carboxylate_dihedral_rb}
  \end{figure}

  \clearpage
% BJC: not necessary
%  \begin{figure}[!htb]
%  \centering
%        \begin{subfigure}{0.45\textwidth}
%                \includegraphics[width=\textwidth]{water_filled_pore.png}
%                \caption{}\label{fig:water_filled_pores}
%        \end{subfigure}
%        \begin{subfigure}{0.45\textwidth}
%                \includegraphics[width=\textwidth]{water_removed.png}
%                \caption{}\label{fig:water_removed}
%        \end{subfigure}
%  \caption{(a) Pores built in the parallel displaced configuration with 5
%	  monomers per layer are filled with 5 wt\% water. (b) The same system is
%	  visualized with water molecules and sodium ions removed. Head groups vacate the
%	  pore region leaving an aqueous solution of water and sodium ions.}\label{fig:water_pores}
%  \end{figure}

@@

  \clearpage
  \section{The role of water in the appearance of R-double}\label{section:rdouble}
